Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2015-09-14 paper.tex
A bunch of work on intro

In this study we set out to model the wake of the University of New Hampshire's
Reference Vertical-Axis (cross-flow) Turbine (UNH-RVAT) using a
Reynolds-averaged Navier--Stokes numerical model.

Note that studies in the literature generally focus on predicting the turbine
loading and the local blade flow rather than the turbine wake. In contrast, we
seek to model the near-wake flow.

It has been shown that the momentum and energy transport processes in the
near-wake of a high-solidity, low aspect ratio vertical-axis turbine are
dominated by vertical advection \cite{Bachant2015-JoT}. It logically follows
that a 2-D simulation would not correctly predict wake recovery, but is of
interest to what extent this is true, since the computational feasibility of 2-D
simulations is attractive.

Modeling the boundary layer flows on cross-flow turbine blades is very
challenging due to the dynamically changing inflow and angle of attack---which
often exceeds static stall values and causes dynamic stall. Furthermore, the
ability to predict the occurrence and interdependence of boundary layer
transition and separation can have dramatic influence on the blade loading and
therefore the predicted turbine power output.

It is assumed that if the turbine power output and near-wake predictions from
the numerical model match those of the experiments, that the flow field can be
inspected in greater detail, i.e., that the experiments will have been
``interpolated''. This will give us access to unmeasured quantities, e.g.
pressure, and allow us to see where the dominant flow structures originate and
help develop low-order wake generator models for use in turbine array modeling.
For example, Araya et al. modeled the flow through a VAT array using leaky
Rankine bodies \cite{Araya2014}. Shamsoddin and Porte-Agel modeled a 2-D CFT
using large eddy simulation (LES) with an actuator line approach to impart force
on the flow field \cite{Shamsoddin2014}.

Goude and Agren used a 2-D vortex method to simulate a farm of cross-flow
turbines \cite{Goude2010}. Durrani et al. used 2-D CFD to model a group of
cross-flow turbines, but did not compare with experimental results
\cite{Durrani2011}.

Questions addressed:

@@

    \item Can 2-D RANS be used for array design?

    \item Does 3-D RANS tell us more about what is happening, i.e, can it
    ``interpolate'' the experimental results?

    \item Does 3-D RANS realize the proper wake recovery proportions? This would
    imply it could be a very accurate model if used in DES.

@@

Goals:

    \item Evaluate modern computational fluid dynamics techniques for predicting
    the dynamic loading on cross-flow turbine blades.

    \item If possible, accurately predict the turbine from the perspective as a
    wake generator, such that the results can be used to develop low-order
    models.

>>>>

Modeling the boundary layer flows on cross-flow turbine blades is very
challenging due to the dynamically changing inflow and angle of attack---which
often exceeds static stall values and causes dynamic stall. Furthermore, the
ability to predict the occurrence and interdependence of boundary layer
transition and separation can have dramatic influence on the blade loading and
therefore the predicted turbine power output. These challenges cast significant
doubt on the ability of high-fidelity CFD modeling to replace wind tunnel or
tank testing of physical models.

\todo[inline]{Put survey of CFD accuracy here}

In this study we set out to model the wake of the University of New Hampshire's
Reference Vertical-Axis (cross-flow) Turbine (UNH-RVAT) using a
Reynolds-averaged Navier--Stokes numerical model. Note that studies in the
literature generally focus on predicting the turbine loading and the local blade
flow rather than the turbine wake. In contrast, we seek to model the near-wake
flow.

The wake of this particular turbine is of interest since it has been shown that
the neark-wake's momentum and energy transport processes are dominated by
vertical advection \cite{Bachant2015-JoT}. It logically follows that a 2-D
simulation would not correctly predict wake recovery, but is of interest to what
extent this is true, since the computational feasibility of 2-D simulations is
attractive.

It is assumed that if the turbine power output and near-wake predictions from
the numerical model match those of the experiments, the flow field can be
inspected in greater detail, i.e., that the experiments will have been
``interpolated''. This will give us access to unmeasured quantities, e.g.
pressure, and allow us to see where the dominant flow structures originate and
help develop low-order wake generator models for use in turbine array modeling.
For example, Shamsoddin and Porte-Agel modeled a 2-D CFT using large eddy
simulation (LES) with an actuator line approach to impart force on the flow
field \cite{Shamsoddin2014}.

To date, little computational work has been done to attempt to design arrays of
CFTs, despite their prospects for closer spacing compared with axial-flow
turbines (AFTs). For example, Araya et al. modeled the flow through a VAT array
using ``leaky Rankine body'' potential flow singularities \cite{Araya2014},
which was able to rank relative performance of array configurations. Goude and
Agren used a 2-D vortex method to simulate a farm of cross-flow turbines
\cite{Goude2010}, though this was not validated with experiments. Durrani et al.
used 2-D CFD to model a group of cross-flow turbines, observing higher power
output for a staggered configuration, but also did not compare with experimental
results \cite{Durrani2011}. Li and Calisal \cite{Li2010} used a 3-D vortex line
method to show mutually improved power output from two adjacent turbines, though
the simulations over-predicted the effects compared with experiments. Antheaume
et al. \cite{Antheaume2008} used a blade element approach coupled with a 3-D
RANS solver to also show how close spacing can improve power output of CFTs.

To summarize, the questions we hope to answer here are:

@@

    \item Can or should 2-D RANS be used for array design?

    \item How accurately can 3-D RANS predict performance?

    \item Can 3-D RANS ``interpolate'' the experimental results and provide insight to develop new low-order wake generators to represent CFTs?

    \item Does 3-D RANS realize the correct proportions of wake recovery
    mechanisms, i.e., are the 3-D blade-resolved results a good ``target'' for
    those from a low-order model?

@@

\section{Turbine description}
